\chapter{Evaluating CRR}

\section*{Using Empirical Results from Research Studies}

\section{Empirical Formula for \(\mathrm{CRR}_{7.5}\)}
The following formulae are derived by accurate determination involving empirical correlations from research on SPT or CPT data:

\begin{equation}
    \mathrm{CRR}_{7.5} = \frac{1}{34 - (\mathrm{N1})_{60}} + \frac{(\mathrm{N1})_{60}}{135} + \frac{50}{(10 \cdot (\mathrm{N1})_{60} + 45)^2} - \frac{1}{200}
\end{equation}

\begin{itemize}
    \item where
    \begin{itemize}
        \item \(\mathrm{CRR}_{7.5}\) is the \(\mathrm{CRR}\) corresponding to a magnitude 7.5 earthquake.
    \end{itemize}
\end{itemize}

\section{Calculation of \(\mathrm{CRR}\)}
Furthermore, \(\mathrm{CRR}\) can be calculated as follows:

\begin{equation}
    \mathrm{CRR} = \mathrm{CRR}_{7.5} \times \mathrm{MSF} \times \mathrm{K}_{\sigma}
\end{equation}

\begin{itemize}
    \item where
    \begin{itemize}
        \item \(\mathrm{CRR}_{7.5}\) is the \(\mathrm{CRR}\) corresponding to a magnitude 7.5 earthquake.
        \item \(\mathrm{MSF}\) is the Magnitude Scaling Factor, which adjusts \(\mathrm{CRR}_{7.5}\) for the actual earthquake magnitude. It can be calculated as follows:
    \end{itemize}
\end{itemize}

\clearpage
\newpage

\section{Magnitude Scaling Factor}
\(\mathrm{MSF}\) is the magnitude scaling factor. It can be calculated as follows:

\begin{equation}
    \mathrm{MSF} = \frac{10^{2.24}}{\mathrm{M}_{w}^{2.56}}
\end{equation}

\begin{itemize}
    \item where
    \begin{itemize}
        \item \(\mathrm{M}_{w}\) is the magnitude of the earthquake.
    \end{itemize}
\end{itemize}

\section{Correction Factor}
\(\mathrm{K}_{\sigma}\) is the correction factor, which adjusts \(\mathrm{CRR}\) for varying overburden stress at different depths or soil conditions. It can be calculated as follows:

\begin{equation}
    \mathrm{K}_{\sigma} = \left(\frac{\sigma_{v0}^{'}}{P_{a}}\right)^{\alpha}
\end{equation}

\begin{itemize}
    \item where
    \begin{itemize}
        \item \(\sigma_{v0}^{'}\) is the effective overburden stress.
        \item \(P_{a}\) is the atmospheric pressure, usually taken to be 100 kPa.
        \item \(\alpha\) is a correction constant, depending on soil type and conditions, usually ranging from 0.4 to 0.7.
    \end{itemize}
\end{itemize}
